Gradual Soundness: Lessons from Static Python

### 2022-04-27 pj-7.1/paper.tex
pj: table 1, remove dict constructor
as Carl says, constructor costs are implied by write costs --- they're free in
typed code, but the constructor column doesn't bring that across

The upside of concrete types is that they enforce stronger type constraints.
These constraints can catch bugs and always lead to faster typed code.
Faster performance is not guaranteed in general, though, because concrete types must
check all writes from untyped contexts.
For comparison, \cref{t:shallow-vs-concrete} presents the worst-case costs of
shallow and concrete dict types.
Shallow types pay a constant-time cost whenever an untyped value enters typed code
and whenever a typed context reads from a shallow value.
Concrete types pay a potentially-high cost for constructors,
a constant-time cost for casts and for writes in untyped code,
and zero cost for reads.
If a value travels to untyped codes and receives many writes but few reads,
then the shallow types may run faster in the end.

@@

  \centering
  \(\begin{array}{l@{}rr@{\qquad}rr@{~~}@{\qquad}r@{~~}r}
                   & \mbox{Constructor} & \mbox{Cast} & \mbox{T-Read} & \mbox{T-Write} & \mbox{U-Read} & \mbox{U-Write} \\\hline
    \sptrawpydict         &   \nocost{} &        O(1) &          O(1) &      \nocost{} &     \nocost{} &      \nocost{} \\
    \sptchkdict{K}{V}     &        O(N) &        O(1) &     \nocost{} &      \nocost{} &     \nocost{} &           O(1)
  \end{array}\)
\end{table}

>>>>

The upside of concrete types is that they enforce stronger type constraints.
These constraints can catch bugs and always lead to faster typed code.
Faster performance is not guaranteed in general, though, because concrete types must
check all writes from untyped contexts.
For comparison, \cref{t:shallow-vs-concrete} presents the worst-case costs of
shallow and concrete dict types.
Shallow types pay a constant-time cost for casts
and whenever a typed context reads from a shallow value.
Concrete types pay nothing for reads, but incur constant-time costs
for casts and for writes from untyped code.
If untyped code frequently writes to a value
(or equivalently, if untyped code initializes a huge value),
then shallow types may run faster.

@@

  \centering
  \(\begin{array}{l@{\qquad}r@{\qquad}rr@{~~}@{\qquad}r@{~~}r}
                                        & \mbox{Cast} & \mbox{T-Read} & \mbox{T-Write} & \mbox{U-Read} & \mbox{U-Write} \\\hline
    \sptrawpydict                       &        O(1) &          O(1) &      \nocost{} &     \nocost{} &      \nocost{} \\
    \sptchkdict{K}{V}                   &        O(1) &     \nocost{} &      \nocost{} &     \nocost{} &           O(1)
  \end{array}\)
\end{table}
